QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-11-23 Revision/Paper_Rev.tex
Rev_V9.2

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a distributed QoE-oriented computation offloading (QECO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. Through numerical studies, we evaluate the performance of QECO. \textcolor{red}{
	Simulation results reveal that compared to the state-of-the-art existing works, QECO increases the number of completed tasks by up to 14.4\%, while simultaneously reducing task delay and energy consumption by 9.2\% and 6.3\%, respectively. Together, these improvements result in a significant average QoE enhancement of 37.1\%. This substantial improvement is achieved by accurately accounting for user dynamics and edge server workloads when making intelligent offloading decisions. This highlights QECO's effectiveness in enhancing users' experience in MEC systems.}

@@

\begin{textblock}{22}(1,0.5)
	\vspace{-9mm}
	\noindent \small \hspace{-8mm} 
	This paper has undergone Revision by IEEE Transactions on Network Science and Engineering (TNSE) on Sep. 4, 2024, [ManuscriptID: TNSE-2024-03-0468]. 
\end{textblock}

@@

To cope with the dynamic nature of the network, recent research has proposed several task offloading algorithms using machine learning methods. In particular, reinforcement learning (RL) \cite{mnih2015human} holds promises to determine optimal decision-making policies by capturing the dynamics of environments and learning strategies for accomplishing long-term objectives. \textcolor{red}{However, the traditional RL methods are not efficient in handling environments with high-dimensional state  spaces. Deep reinforcement learning (DRL) combines  traditional RL  with  deep  neural  networks  to  intelligently respond to the unknown and dynamic system when addressing the limitations of RL-based algorithms. Despite these advancements, task offloading still faces significant challenges in real-world scenarios with multiple MDs and ENs. In such scenarios, it is essential for MDs to make offloading decisions independently  without prior knowledge of other MDs’ tasks and offloading models. However,  existing works fail to adequately address this challenge. In addition, QoE is a time-varying performance measure that reflects user satisfaction and is not affected only by delay \cite{guo2022energy}, but also by energy consumption \cite{tang2022uav}. Albeit some existing works have investigated the trade-off between delay and energy consumption \cite{li2022joint}, they fail to properly estimate the QoE value and address the users' requirements. Although deep Q-networks (DQN) offer a valuable tool, a more comprehensive approach is required to accurately estimate the QoE while addressing the aforementioned challenges in practical scenarios where global information is limited.}

In this study, we delve into the computation task offloading problem in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance.   \textcolor{red}{We propose a distributed QoE-oriented computation offloading (QECO) algorithm to efficiently handle task offloading under uncertain loads at ENs.} This algorithm empowers MDs to make offloading decisions utilizing only locally observed information, such as task size, queue details, battery status, and historical workloads at the ENs. 
  \textcolor{blue}{To capture the dynamic nature of the MEC environment, we employ the dueling double deep Q-network (D3QN), which is a refined improvement over standard DQN model. By integrating both double Q-learning \cite{van2016deep} and dueling network architectures \cite{wang2016dueling}, D3QN reduces overestimation bias in action-value predictions and more accurately distinguishes the relative importance of states and actions. We also incorporate long short-term memory (LSTM) \cite{hochreiter1997long} into our D3QN model to continuously estimate dynamic workloads at ENs. This enables MDs to effectively handle the uncertainty of the MEC environment, where global information is limited, and proactively adjust the offloading strategies to improve their long-term QoE estimation.} By adopting the appropriate policy based on each MD’s specific requirements at any given time, the QECO algorithm significantly improves the QoE for individual users.

@@

\begin{itemize}
	\item \textit{Task Offloading Problem in the MEC System:} \textcolor{red}{ We consider queuing systems for MDs and ENs and formulate the task offloading problem taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). We aim to maximize the long-term QoE, which reflects the task completion rate, task delay, and energy consumption of MDs. Our problem formulation effectively utilizes the resources and properly handles the dynamic fluctuations of workload at ENs. }
	%We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). Given task completion, task delay, and energy consumption as the QoE requirements of MDs, this approach concentrates on maximizing the long-term QoE, by effectively utilizing the resources and properly handling the dynamic nature of workload at ENs.

@@

	\item \textit{Performance Evaluation:} \textcolor{red}{We conduct comprehensive experiments to evaluate QECO’s performance and its training convergence against baseline schemes as well as two state-of-the-art existing works. The results demonstrate that our algorithm effectively utilizes the computing resources while taking the dynamic workloads at ENs into consideration. In particular, QECO converges more quickly than vanilla DQN and DDQN methods. QECO also improves average QoE by 37.1\% and 29.8\% compared to the potential game-based offloading algorithm (PGOA)~\cite{yang2018distributed} and distributed and collective DRL-based offloading algorithm (DCDRL)~\cite{qiu2020distributed}, respectively.}

@@

\color{black}
\section{Related Work}
\label{section:II}

@@

For time-sensitive applications, minimizing task delay is the primary optimization objective, driving the development of delay-optimal approaches in MEC. 
To solve the delay minimization problem in MEC environment, Zhang \textit{et al.} in \cite{zhang2023offline} proposed an offline RL-based computation offloading algorithm, where the problem is modeled as a repeated game between two agents.
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. Liu \textit{et al.} in \cite{liu2022deep} formulated the problem as MDP and proposed DRL-based approaches to minimaze both communication and computation delay in MEC.
To minimize total delay and reduce mobile vehicle task waiting time, Tang \textit{et al.} in \cite{tang2022double} developed a dynamic offloading model for multiple mobile vehicles, segmenting tasks into sequential subtasks for more precise offloading decisions.
 Li \textit{et al.} in \cite{li2022integrated} proposed an integrated optimization framework, which leverages past experience and model knowledge to enable fast and resilient real-time offloading control. 
In \cite{wei2023many}, the authors investigated deadline-constraiend for time-sensetive tasks in vehicular MEC network and proposed a multi agent RL-based computation offloading algorithm based on POMDP to addresses jointly delay and computation rate optimazation. However, delay-optimal approaches~\cite{zhang2023offline}--\cite{wei2023many} often sacrifice other key performance metrics and may cause greater energy consumption or reduce the overall computation capacity, to achieve their goal.  
  %\cite{zheng2023drl} \cite{xue2023joint}

For energy-constrained MEC systems, reducing energy consumption becomes a key objective, motivating the creation of energy-efficient approaches. 
Munir \textit{et al.} \cite{munir2021multi}  investigated an energy dispatch problem of a self-powered wireless network with MEC capabilities and developed a semi-distributed approach using a multi-agent RL framework to minimize energy consumption.
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to optimize resource allocation and computation offloading concurrently. 
%%In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
To ensure that minimizing energy consumption does not lead to a reduction in the system's computing rate, researches have expanded their work by incorporating the maximum tolerable delay for task models.
In \cite{chouikhi2023energy}, Chouikhi \textit{et al.} proposed a computation offloading approach based on DRL to minimize long-term energy consumption and maximize the number of tasks completed before their tolerant deadlines.
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network. 
Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
However, energy-efficient approaches~\cite{munir2021multi}--\cite{huang2021deadline}, even when considering deadline constraints, do not adequately address the complexities of time-sensitive applications, where reducing delay can have significant effects on the user's QoE. Particularly, saving transmission energy consumption may reduce the transmission power of MDs and lead to poor performance for real-time applications.

To address time-sensitive applications in energy-constrained MEC systems, solving the relationship between delay and energy consumption is the primary optimization objective.
Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments. 
 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 Liu \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. 
 She \textit{et al.} in \cite{she2024efficient} investigated the computation offloading optimization problem and proposed DRL-based approach to minimize both the transmission delay and energy, based on POMDP.
  Although delay-energy trade-off approaches~\cite{liao2023online}--\cite{she2024efficient} effectively address time-sensitive tasks, they are not suitable for applications with strict requirements on delay or energy consumption.

Incorporating task processing deadlines for time-sensitive applications in energy-constrained MEC systems requires achieving an effective balance among delay, energy consumption, and computation rate. 
  In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.
 To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent DQN-based algorithm. 
  In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
  Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
  Wu \textit{et al.} in \cite{wu2023computation} introduced a stochastic game-based resource allocation in the SDN-based MEC network. They used an MDP and proposed a multi-agent RL method to minimize both energy consumption and processing delay. While studies \cite{gao2022large}--\cite{wu2023computation} have explored delay-sensitive applications with prossecing deadline and addressed energy-constrained MEC, they overlook user-specific demands and fail to meet QoE requirements effectively.

In contrast to aformentationed related works \cite{zhang2023offline}--\cite{wu2023computation}, which overlook key performance metrics, we are motivated to effectively address QoE optimization, where QoE accounts for task completion rate, processing delay, and energy consumption. Different from \cite{zhang2023offline}--\cite{li2022integrated}, \cite{munir2021multi}--\cite{dai2020edge}, and \cite{liao2023online}--\cite{she2024efficient}, which considered delay-tolerant tasks, we take into account delay-sensitive tasks with processing deadlines. This is challenging to address, as processing deadlines impact the load dynamics at the ENs, which in turn affect the delay of offloaded tasks. 
  To effectively address personalized QoE optimization, we designed an adaptive balance for these factors according to the user's preferences, such as prioritizing reduced task delay or energy saving, which may vary over time. Each user may have unique QoE requirements at any given time.

QoE is a time-varying performance measure that reflects user satisfaction, where each user may have unique QoE requirements. According to the MDs' individual preferences, each delay, energy consumption, and computation rate may take precedence over two others.

In addition, the limited processing capacities of ENs necessitate careful consideration when allocating resources to MDs. The processing capacity allocated to an MD depends on the dynamic workload of an EN at any given time. Some studies have failed to account for these challenges in dynamic network environments, resulting in reduced system performance. For example, in \cite{chen2021drl}, it is assumed that MDs have sufficient transmission and computation resources from ENs, which may not reflect real-world conditions. While the works \cite{zhao2019deep}, \cite{dai2020edge}, and \cite{wu2024privacy} investigated limited resources in MEC, they did not consider dynamic workloads at ENs in decision-making processes, leading to potential performance issues.
Certain studies such as \cite{sun2024hierarchical}, \cite{gong2022edge}-\cite{gao2022large} have addressed load levels at ENs and proposed task offloading algorithms. However, these algorithms often require global system information, resulting in high signaling overhead. In contrast to these works, we propose a distributed algorithm that effectively manages unknown load dynamics at ENs. This algorithm enables each MD to make offloading decisions independently, without requiring information (e.g., task models, offloading decisions) from other MDs, thereby reducing signaling overhead and improving overall system performance.

%Some existing works primarily focus on simplified MEC networks, investigating single MEC network models. For instance, studies \cite{zhou2021deep}, \cite{liu2021learn} only examine single-server or single-user scenarios, while \cite{huang2019deep}--\cite{wu2023computation} do not account for delay-sensitive tasks. In contrast, our research addresses a more realistic MEC scenario, considering delay-sensitive tasks with processing deadlines. This is challenging due to the interplay between processing deadlines, load dynamics at the ENs, and the resulting offloaded task delays. Additionally, the absence of robust underlying resource scheduling systems can lead to data congestion during transmission and inefficient resource allocation, hindering overall system performance \cite{wang2020reinforcement}--\cite{chen2021drl}. Hence, more accurate MEC models are necessary to handle the complexities of dynamic resource scheduling in large-scale scenarios.
%To better represent real-world scenarios, authors in \cite{Bolourian-WCL24}--\cite{9253665} have applied queuing theory to model MEC systems. This approach helps capture the dynamic nature of networks, enabling MEC systems to allocate computational and transmission resources more efficiently. In a similar vein, we approached the offloading process as a holistic queuing system, where multiple queues ensure fair resource scheduling for MEC, leading to more accurate and efficient resource management.

 \begin{table*}[htbp]\textcolor{blue}{
 		\renewcommand{\arraystretch}{0.2}

@@

 		\scalebox{1}{%
 			\begin{tabular}{ lp{5.4cm}p{4cm}p{2cm}p{4cm}l} 

@@

 				\textbf{Paper}  &  \textbf{Problem} &  \textbf{Objective} & \textbf{Method} &  \textbf{Limitation}  \\ \midrule
 				\cite{li2022integrated} & Online computation offloading & Minimize computation delay & Actor-critic network & Overlooking energy consumtion \\ \midrule
 				%\cite{li2020joint} & Caching and computation offloading & Minimize total system delay &  Multi-agent DQN & Not consider system energy consumtion \\ \midrule
 				\cite{tang2022double}  & Computation offloading & Minimize total system delay & DDQN & Overlooking energy consumtion \\ \midrule
 				%\cite{wang2020reinforcement} & Mobility aware computation offloading & Minimize energy consumtion & Q-learning & Not consider computation delay\\ \midrule
 				\cite{liu2021learn}  & Resource allocation and computation offloading & Minimize computation cost  & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
 				\cite{wu2023computation}   & Stochastic game-based resource allocation & Minimize computation delay and energy
consumption & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
 				\cite{zhou2021deep} & Resource allocation and computation offloading & Minimize total energy consumption & DDQN & Limited to a single edge server \\\midrule				
 				%\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
 				\cite{wu2024privacy} &  power allocation and computation offloading  & Optimize quality of service  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
 				\cite{dai2020edge} & Resource allocation and computation offloading & Minimize energy consumption & DDPG & Not take user's demand into consideration \\\midrule
 				\cite{zhao2019deep} & User association and resource allocation & Optimize downlink utility  & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
 				\cite{liao2023online}  & Online computation offloading & Minimize computation delay and energy consumption & DDQN & Only for a single MEC system model  \\\midrule
 				\cite{wu2024combining}   & Privacy aware computation offloading & Optimization of computation rate and energy consumption  & Actor-critic network  & Not take user's demand into consideration \\\midrule
 				\cite{Bolourian-WCL24}  & Energy harvesting  computation offloading & Minimize drop rate and energy consumption.  & DQN & Does not account for computation delay  \\\midrule
 				\cite{huang2021deadline}  & Computation offloading & Minimize drop rate and energy consumption & DDPG & Not consider computation delay \\ \midrule
 				\cite{gong2022edge} & Resource allocation and computation offloading & Minimize computation delay and energy
consumption & DRL & Not take user's demand into consideration\\\midrule
 				\cite{sun2024hierarchical} & Computation offloading and service caching & Minimize average delay & Hierarchical DRL & Not consider system energy consumtion \\\midrule
 				\cite{wu2023multi}  & Computation offloading & Minimize computation delay and energy
consumption  & Multi-agent PPO&  Not consider tasks with maximum delay tolerance \\\midrule
 				\cite{9253665}  &  Computation offloading & Minimize drop rate and computation delay.  & D3QN + LSTM & Not consider system energy consumtion\\\midrule
 				\cite{gao2022large}  & Decentralized computation offloading & Optimize system cost and computation rates  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
 				QECO  &  Computation offloading & Maximize the QoE of each MD individually & D3QN + LSTM & \\

@@

%\subsection{Analysis and Motivation}
%Despite extensive research efforts, task offloading continues to encounter significant challenges in real-world scenarios involving multiple MDs and ENs. We discuss these challenges from two aspects, i.e., system modeling and target objective optimization.

 % \subsubsection{System modeling}

%Some existing works primarily focus on simplified MEC networks, investigating single MEC network models. For instance, studies \cite{zhou2021deep}, \cite{liu2021learn} only examine single-server or single-user scenarios, while \cite{huang2019deep}--\cite{wu2023computation} do not account for delay-sensitive tasks. In contrast, our research addresses a more realistic MEC scenario, considering delay-sensitive tasks with processing deadlines. This is challenging due to the interplay between processing deadlines, load dynamics at the ENs, and the resulting offloaded task delays. Additionally, the absence of robust underlying resource scheduling systems can lead to data congestion during transmission and inefficient resource allocation, hindering overall system performance \cite{wang2020reinforcement}--\cite{chen2021drl}. Hence, more accurate MEC models are necessary to handle the complexities of dynamic resource scheduling in large-scale scenarios.
%To better represent real-world scenarios, authors in \cite{Bolourian-WCL24}--\cite{9253665} have applied queuing theory to model MEC systems. This approach helps capture the dynamic nature of networks, enabling MEC systems to allocate computational and transmission resources more efficiently. In a similar vein, we approached the offloading process as a holistic queuing system, where multiple queues ensure fair resource scheduling for MEC, leading to more accurate and efficient resource management.

%In addition, the limited processing capacities of ENs necessitate careful consideration when allocating resources to MDs. The processing capacity allocated to an MD depends on the dynamic workload of an EN at any given time. Some studies have failed to account for these challenges in dynamic network environments, resulting in reduced system performance. For example, in \cite{chen2021drl}, it is assumed that MDs have sufficient transmission and computation resources from ENs, which may not reflect real-world conditions. While the works \cite{zhao2019deep}, \cite{dai2020edge}, and \cite{wu2024privacy} investigated limited resources in MEC, they did not consider dynamic workloads at ENs in decision-making processes, leading to potential performance issues.
%Certain studies such as \cite{sun2024hierarchical}, \cite{gong2022edge}-\cite{gao2022large} have addressed load levels at ENs and proposed task offloading algorithms. However, these algorithms often require global system information, resulting in high signaling overhead. In contrast to these works, we propose a distributed algorithm that effectively manages unknown load dynamics at ENs. This algorithm enables each MD to make offloading decisions independently, without requiring information (e.g., task models, offloading decisions) from other MDs, thereby reducing signaling overhead and improving overall system performance.

>>>>

\usepackage{multirow}

\renewcommand{\arraystretch}{1.2}

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a distributed QoE-oriented computation offloading (QECO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. Through numerical studies, we evaluate the performance of QECO. \textcolor{blue}{
	Simulation results reveal that compared to the state-of-the-art existing works, QECO increases the number of completed tasks by up to 14.4\%, while simultaneously reducing task delay and energy consumption by 9.2\% and 6.3\%, respectively. Together, these improvements result in a significant average QoE enhancement of 37.1\%. This substantial improvement is achieved by accurately accounting for user dynamics and edge server workloads when making intelligent offloading decisions. This highlights QECO's effectiveness in enhancing users' experience in MEC systems.}

@@

%\begin{textblock}{22}(1,0.5)
%	\vspace{-9mm}
%	\noindent \small \hspace{-8mm} 
%	This paper has undergone Revision by IEEE Transactions on Network Science and Engineering (TNSE) on Sep. 4, 2024, [ManuscriptID: TNSE-2024-03-0468]. 
%\end{textblock}

@@

To cope with the dynamic nature of the network, recent research has proposed several task offloading algorithms using machine learning methods. In particular, reinforcement learning (RL) \cite{mnih2015human} holds promises to determine optimal decision-making policies by capturing the dynamics of environments and learning strategies for accomplishing long-term objectives. \textcolor{blue}{However, the traditional RL methods are not efficient in handling environments with high-dimensional state  spaces. Deep reinforcement learning (DRL) combines  traditional RL  with  deep  neural  networks  to  intelligently respond to the unknown and dynamic system when addressing the limitations of RL-based algorithms. Despite these advancements, task offloading still faces significant challenges in real-world scenarios with multiple MDs and ENs. In such scenarios, it is essential for MDs to make offloading decisions independently  without prior knowledge of other MDs’ tasks and offloading models. However, existing works fail to adequately address this challenge. In addition, QoE is a time-varying performance measure that reflects user satisfaction and is not affected only by delay, but also by energy consumption. Albeit some existing works have investigated the trade-off between delay and energy consumption, they fail to properly estimate the QoE value and address the users' requirements. Although deep Q-networks (DQN) offer a valuable tool, a more comprehensive approach is required to accurately estimate the QoE while addressing the aforementioned challenges in practical scenarios where global information is limited.}
%delay \cite{guo2022energy}
%energy consumption \cite{tang2022uav}
%trade-off between delay and energy consumption \cite{li2022joint}

In this study, we delve into the computation task offloading problem in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance.   \textcolor{blue}{We propose a distributed QoE-oriented computation offloading (QECO) algorithm to efficiently handle task offloading under uncertain loads at ENs.} This algorithm empowers MDs to make offloading decisions utilizing only locally observed information, such as task size, queue details, battery status, and historical workloads at the ENs. 
  \textcolor{blue}{To capture the dynamic nature of the MEC environment, we employ the dueling double deep Q-network (D3QN), which is a refined improvement over standard DQN model. By integrating both double Q-learning \cite{van2016deep} and dueling network architectures \cite{wang2016dueling}, D3QN reduces overestimation bias in action-value predictions and more accurately distinguishes the relative importance of states and actions. We also incorporate long short-term memory (LSTM) \cite{hochreiter1997long} into our D3QN model to continuously estimate dynamic workloads at ENs. This enables MDs to effectively handle the uncertainty of the MEC environment, where global information is limited, and proactively adjust the offloading strategies to improve their long-term QoE estimation.} By adopting the appropriate policy based on each MD’s specific requirements at any given time, the QECO algorithm significantly improves the QoE for individual users.

@@

\begin{itemize}
	\item \textit{Task Offloading Problem in the MEC System:} \textcolor{blue}{ We consider queuing systems for MDs and ENs and formulate the task offloading problem taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). We aim to maximize the long-term QoE, which reflects the task completion rate, task delay, and energy consumption of MDs. Our problem formulation effectively utilizes the resources and properly handles the dynamic fluctuations of workload at ENs. }
	%We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). Given task completion, task delay, and energy consumption as the QoE requirements of MDs, this approach concentrates on maximizing the long-term QoE, by effectively utilizing the resources and properly handling the dynamic nature of workload at ENs.

@@

	\item \textit{Performance Evaluation:} \textcolor{blue}{We conduct comprehensive experiments to evaluate QECO’s performance and its training convergence against baseline schemes as well as two state-of-the-art existing works. The results demonstrate that our algorithm effectively utilizes the computing resources while taking the dynamic workloads at ENs into consideration. In particular, QECO converges more quickly than vanilla DQN and DDQN methods. QECO also improves average QoE by 37.1\% and 29.8\% compared to the potential game-based offloading algorithm (PGOA)~\cite{yang2018distributed} and distributed and collective DRL-based offloading algorithm (DCDRL)~\cite{qiu2020distributed}, respectively.}

@@

\color{blue}
\section{Related Work}
\label{section:II}

@@

\begin{table*}[tbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{1}

@@

		\scalebox{1.05}{
			\begin{tabular}{ lp{4.4cm}p{0.5cm}p{0.6cm}p{0.4cm}p{0.8cm}p{2.2cm}p{3.7cm}l} 

@@

				\multirow{2}{*}{{\textbf{Paper}}}&\multirow{2}{*}{{\textbf{Problem}}} &\multicolumn{4}{@{}l@{}}{\textbf{Optimization target}}&\multirow{2}{*}{{\textbf{Method}}} & \multirow{2}{*}{{\textbf{Limitation}}}   \\ \cmidrule{3-6}
				&&Delay&Energy& Drop& QoE &&&\vspace{1mm}  \\ 							\toprule
				\cite{zhang2023offline}  & Computation offloading &  \centering \cmark & \centering \xmark&  \centering \xmark&  \centering \xmark& Offline RL & \multirow{6}{*}{\shortstack[l]{Overlooking energy consumption; \\Not suitable for energy-constrained\\ MEC; May cause greater energy \\consumption or reduce the overall\\ system capacity.}} \\ \cmidrule{1-7}
				\cite{sun2024hierarchical}  & Joint offloading and service caching & \centering \cmark & \centering \xmark&  \centering \xmark&  \centering \xmark& Hierarchical DRL &   \\ \cmidrule{1-7}%\midrule
				\cite{liu2022deep} & Computation offloading  &  \centering \cmark &  \centering \xmark &  \centering \xmark&  \centering \xmark& DQN, DDPG&    \\  \cmidrule{1-7}
				\cite{tang2022double} & Computation offloading &  \centering \cmark & \centering \xmark&  \centering \xmark&  \centering \xmark& DDQN &   \\  \cmidrule{1-7}
				\cite{li2022integrated} &  Online computation offloading &  \centering \cmark & \centering \xmark&  \centering \xmark&  \centering \xmark& AC network&   \\  \cmidrule{1-7}
				\cite{wei2023many} & Computation offloading  &  \centering \cmark & \centering \xmark&  \centering \cmark&\centering\xmark & AC network  &&\vspace{1mm}  \\  								\toprule
				\cite{munir2021multi}  & Computation offloading & \centering\xmark &\centering\cmark&\centering \xmark& \centering\xmark& Multi-agent RL & \multirow{5}{*}{\shortstack[l]{Do not account for time-sensetive\\ applications; Saving transmission \\energy consumption may reduce \\the transmission power of MDs.}}  \\ \cmidrule{1-7}
				\cite{zhou2021deep}  & Joint offloading and resource allocation & \centering\xmark &\centering\cmark& \centering\xmark& \centering\xmark& DDQN &   \\ \cmidrule{1-7}%\midrule
				\cite{dai2020edge} & Joint offloading and resource allocation & \centering\xmark &\centering\cmark& \centering\xmark&\centering\xmark& DDPG &   \\  \cmidrule{1-7}
				\cite{chouikhi2023energy} & Computation offloading & \centering\xmark &\centering\cmark&\centering \cmark&\centering\xmark& DRL &   \\  \cmidrule{1-7}
				\cite{wu2024combining} & Privacy aware computation offloading &\centering \xmark &\centering\cmark&\centering \cmark&\centering\xmark&AC network && \vspace{1mm}  \\  	\toprule
				\cite{huang2021deadline} & Computation offloading  & \centering\cmark &\centering\cmark& \centering\xmark&\centering\xmark& DDPG&\multirow{6}{*}{\shortstack[l]{Not suitable for applications with \\strict delay requirements, as \\processing deadlines influence the \\load dynamics at the ENs, impacting \\the delay of offloaded tasks. }} \\  			 \cmidrule{1-7}
				\cite{liao2023online} & Online computation offloading & \centering\cmark &\centering\cmark& \centering\xmark& \centering\xmark&DDQN& \\  			 \cmidrule{1-7}
				\cite{wu2023multi} & Computation offloading & \centering\cmark &\centering\cmark& \centering\xmark&\centering\xmark& Multi-agent PPO&  \\  			 \cmidrule{1-7}
				\cite{gong2022edge} & Joint offloading and resource allocation & \centering\cmark &\centering\cmark& \centering\xmark&\centering\xmark& DRL &   \\  			 \cmidrule{1-7}
				\cite{liu2021learn}  &  Joint offloading and resource allocation  & \centering\cmark &\centering\cmark& \centering\xmark&\centering\xmark & Multi-agent AC &   \\  \cmidrule{1-7}
				\cite{she2024efficient} & Computation offloading & \centering\cmark &\centering\cmark& \centering\xmark&\centering\xmark& Multi-agent DDPG &&\vspace{1mm}    \\  						\toprule
				\cite{gao2022large} & Computation offloading & \centering\cmark &\centering\cmark& \centering\cmark& \centering\xmark& Multi-agent AC&\multirow{5}{*}{\shortstack[l]{Overlooking personalized QoE; \\Treating all MDs uniformly in\\ performance optimization fails to \\address the user satisfaction.}}  \\  			 \cmidrule{1-7}
				\cite{wu2024privacy} &Joint offloading and power allocation & \centering\cmark &\centering\cmark& \centering\cmark&\centering\xmark& Multi-agent AC  &   \\  			 \cmidrule{1-7}
				\cite{Bolourian-WCL24} & Energy harvesting offloading & \centering\cmark &\centering\cmark& \centering\cmark&\centering\xmark& DQN&   \\  			 \cmidrule{1-7}
				\cite{chen2021drl} & Joint offloading and resource allocation& \centering\cmark &\centering\cmark& \centering\cmark&\centering\xmark& DDPG &   \\  			 \cmidrule{1-7}
				\cite{wu2023computation} &Stochastic game resource allocation & \centering\cmark &\centering\cmark&\centering\cmark& \centering\xmark& Multi-agent RL & &\vspace{1mm}  \\  										\toprule
				QECO & QoE-oriented computation offloading  & \centering\cmark &\centering\cmark& \centering\cmark&\centering\cmark& D3QN + LSTM &&\vspace{1mm} 	 		  \\

@@

For time-sensitive applications, minimizing task delay is the primary optimization objective, driving the development of delay-optimal approaches in MEC. 
To solve the delay minimization problem in MEC environment, Zhang \textit{et al.} in \cite{zhang2023offline} proposed an offline RL-based computation offloading algorithm, where the problem is modeled as a repeated game between two agents.
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. Liu \textit{et al.} in \cite{liu2022deep} formulated the problem as MDP and proposed DRL-based approaches to minimize both communication and computation delay in MEC.
To minimize total delay and reduce mobile vehicle task waiting time, Tang \textit{et al.} in \cite{tang2022double} developed a dynamic offloading model for multiple mobile vehicles, segmenting tasks into sequential subtasks for more precise offloading decisions.
Li \textit{et al.} in \cite{li2022integrated} proposed an integrated optimization framework, which leverages past experience and model knowledge to enable fast and resilient real-time offloading control. 
In \cite{wei2023many}, the authors investigated deadline-constrained time-sensitive tasks in vehicular MEC networks and proposed a multi-agent RL-based computation offloading algorithm based on POMDP to address joint delay and computation rate optimization. However, delay-optimal approaches~\cite{zhang2023offline}--\cite{wei2023many} often sacrifice other key performance metrics and may cause greater energy consumption or reduce the overall computation capacity, to achieve their~goal.
